---
title: "HW 4"
author: "Hollie Pennington"
date: '2022-05-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rootSolve)
library(janitor)
library(here)
```

```{r}
#read in data 
damages_df <- read_csv(here("data","damages.csv")) %>% 
  clean_names()
warming_df <- read_csv(here("data","warming.csv")) %>% 
  clean_names()
```

ESM 204 HW 4:

Calculating the SCC and policy choice under uncertainty.

In this homework, you will use new estimates of the future impacts of climate change to inform an update to the SCC. We recommend using R and writing functions to compute your answers wherever possible.

Gauchospace houses two datasets. The first, “damages.csv” provides new model estimates of the annual total damages from climate change at different levels of warming (in degrees C). 
The second dataset, “warming.csv” contains estimates of a baseline future climate trajectory (in degrees C) until 2100, and a second trajectory that adds a one-time pulse of CO2 today to the atmosphere. The pulse is 35 billion tons of carbon, which is roughly equal to annual global emissions. You can think of this as a “small” one-time pulse in carbon emissions.

# 1. Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Estimate a model with a zero intercept because damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

```{r}
#quadratic damage function 
# the -1 is from Sandy's OH
# then use predict function
d_quad_model <- lm(damages ~ warming + I(warming^2) -1, data = damages_df) 
d_predicted <- d_quad_model$fitted.values
```

```{r}
# function and scatter plot graph for damages
ggplot(d_quad_model, aes(warming, damages)) + 
  geom_line(aes(y = d_predicted)) + 
  geom_point()
```

# 2. Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: 

## (1) damages over time without the pulse

```{r}
#maybe we should merge data sets? 
warming <- rename(warming_df, warming = warming_baseline)

merged_df <- left_join(warming, damages_df, by = "warming")
#this did not work 
```

```{r}
#Use warming data and predict model 
#not used for the graph but I am keeping it encase I need to make one later
w_quad_model <- lm(year ~ warming_baseline + I(warming_baseline^2) -1, data = warming_df) 
w_predicted <- w_quad_model$fitted.values
```

```{r}
#create plot 
ggplot(warming_df, aes(year, warming)) + 
  geom_point() + 
  geom_line(data = damages_df, aes(d_predicted))
#  I need to find a way to add this d_predicted function to the graph 
```

## (2) damages over time with the pulse

## (3) the difference in damages over time that arises from the pulse

## (4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).

# 3. The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC.

```{r}

```

# 4. The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation: r = ρ + ηg Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

# 5. Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual and Policy B is to take immediate and strong action on climate change. Use these facts

• If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

In principle, we would like the “pulse” to be a single ton of carbon. But this pulse hardly shows up unless you measure temperature out to 10 decimal points. So we introduce a larger one-time pulse and you can infer the “per-ton” effects by dividing the effect by 35 billion.

• Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever.
• Society is risk neutral
• Use a discount rate of 2%
## What is the expected present value of damages up to 2100 under Policy A? 

## What is the expected present value of damages up to 2100 under Policy B? 

## Suppose undertaking Policy A costs zero and undertaking Policy B costs X. The total cost of a policy is the implementation cost plus expected damages. Make a plot of total cost of Policy A and total cost of Policy B against X. 

## Suppose Congress wants to undertake the policy with the lower total cost. Use your graph to advise Congress.

# 6. Now suppose that society is risk averse. How do you think this would change your advice to Congress?
