---
title: "HW 4"
author: "Grace Bianchi, Andrea Cornelius, Hollie Pennington"
date: '2022-05-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(rootSolve)
library(janitor)
library(here)
```

```{r}
#read in data 
damages_df <- read_csv(here("data","damages.csv")) %>% 
  clean_names()
warming_df <- read_csv(here("data","warming.csv")) %>% 
  clean_names()
```
```{r}
#quadratic damage function 
# the -1 is from Sandy's OH it is to make the model start at zero
# then use predict function
d_quad_model <- lm(damages ~ warming + I(warming^2) -1, data = damages_df) 
d_predicted <- d_quad_model$fitted.values
d_quad_model
```

```{r}
# function and scatter plot graph for damages
ggplot(d_quad_model, aes(warming, damages)) + 
  geom_line(aes(y = d_predicted)) + 
  geom_point() +
  labs(x= "Global warming (degrees C)", y = "Damages ($)") +
  theme_minimal()
```

# 2. Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: 

```{r}
#office hour notes: 
#use coefficients from d_quad and multiply by warming baseline  
#use mutate to create damages column in warming.csv
#do the same thing when it is time to do the pulse 
#create damages function 
dam_function <-function(warming)
{dam <- (d_quad_model$coefficients[2]) *(warming^2) + 
  d_quad_model$coefficients[1] * (warming)
return(dam)}
#calculate warming damages for baseline and pulse 
mutated_df <- warming_df %>% 
  mutate(dam_baseline = dam_function(warming_baseline)) %>% 
    mutate(dam_pulse = dam_function(warming_pulse)) %>% 
    mutate(dam_add = (dam_pulse - dam_baseline)) %>% 
    mutate(ton_dam = (dam_add/35000000000))
#the mutated_df has the additional columns created from running the damage function
#dam_add is the additional damages caused by the pulse for 2(3).
#ton_dam is the conversion for 2(4)
```

## (1) damages over time without the pulse (Baseline damage)

```{r}
#Use warming data and predict model 
#not used for the graph but I am keeping it in case I need to make one later
w_quad_model <- lm(year ~ warming_baseline + I(warming_baseline^2) -1, data = warming_df) 
w_predicted <- w_quad_model$fitted.values
```

```{r}
#create plot 
ggplot(mutated_df, aes(x= year, y = dam_baseline)) + 
  geom_point() + 
  geom_line() +
  labs(x = "Year", y =  "Damages ($)", title = "Baseline Damages (without pulse)")
```

## (2) damages over time with the pulse


```{r}
#create plot 
ggplot(mutated_df, aes(x= year, y = dam_pulse)) + 
  geom_point() + 
  geom_line() +
  labs(x = "Year", y =  "Damages($)", title = "Damages with pulse")
```

## (3) the difference in damages over time that arises from the pulse
```{r}
#use dam_add in mutated_df to plot 
ggplot(mutated_df, aes(x = year, y = dam_add)) + 
  geom_line() + 
  geom_point() +
   labs(x = "Year", y =  "Damage Added (Tons of CO2)", title = "Additional damages with pulse")
```

## (4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).

```{r}
#do a conversion from $damages/35_billion_tons_co2 to $damages/ton_CO2 -- Hollie completed in earlier code chunk as ton_dam
ggplot(mutated_df, aes(x = year, y = ton_dam)) + 
  geom_line() + 
  geom_point() +
  labs(x = "Year", y =  "Damages per ton of CO2 (USD)", title = "Damages with pulse per ton of CO2")
```

# 3. The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC.

```{r}
#office hour notes
#the scc is not the baseline -- it is the marginal 
#Scc = present discounted value 
#scc = summation: T=2100(top) Epsilon t=2022(bottom)(damages from pulse^t)/(1+r)^t
#then plot r (x axis) and scc (y axis)

scc <- mutated_df %>% 
  mutate(years_into_future = year-2022) %>% 
  mutate(scc_not_discounted = dam_add/35e+9) %>% 
  mutate(dr0.25 = scc_not_discounted/(1.0025^years_into_future)) %>% 
  mutate(dr0.5 = scc_not_discounted/(1.005^years_into_future)) %>% 
  mutate(dr0.75 = scc_not_discounted/(1.0075^years_into_future)) %>% 
  mutate(dr1 = scc_not_discounted/(1.01^years_into_future)) %>% 
  mutate(dr1.5 = scc_not_discounted/(1.015^years_into_future)) %>% 
  mutate(dr2 = scc_not_discounted/(1.02^years_into_future)) %>% 
  mutate(dr3 = scc_not_discounted/(1.03^years_into_future)) %>% 
  mutate(dr4 = scc_not_discounted/(1.04^years_into_future)) %>% 
  mutate(dr5 = scc_not_discounted/(1.05^years_into_future)) %>% 
  mutate(dr6 = scc_not_discounted/(1.06^years_into_future)) %>% 
  mutate(dr2.1 = scc_not_discounted/(1.021^years_into_future))

scc0.025 <- sum(scc$dr0.25)
scc0.5 <- sum(scc$dr0.5)
scc0.75 <- sum(scc$dr0.75)
scc1 <- sum(scc$dr1)
scc1.5 <- sum(scc$dr1.5)
scc2 <- sum(scc$dr2)
scc3 <- sum(scc$dr3)
scc4 <- sum(scc$dr4)
scc5 <- sum(scc$dr5)
scc6 <- sum(scc$dr6)

scc_vector <- c(scc0.025, scc0.5, scc0.75, scc1, scc1.5, scc2, scc3, scc4, scc5, scc6)
dr <- c(0.25, 0.5, 0.75, 1, 1.5, 2, 3, 4, 5, 6)
scc_table <- data.frame(scc_vector, dr)
```

```{r}
ggplot(data = scc_table, aes(x = dr, y = scc_vector)) +
    geom_line() +
  labs(x = "Discount rate (%)", y = "Social cost of carbon") +
  theme_minimal() 
```
Discount rate is the value placed on near-term benefits and damages versus future benefits and damages. As the discount rate increases, more value is placed on near-term benefits and damages than future benefits and damages, and vice versa.The social cost of carbon reflects the price we set for future damages of one ton of carbon emissions. So, as the discount rate increases, we will weigh future damages as less important than near-term damages and benefits, and the social cost of carbon will be lower.

# 4. The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation: r = ρ + ηg Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

```{r}
r <- (0.001 + 2*0.01) * 100
```

 **The discount rate is 2.1% using the Ramsey rule.**

```{r}
scc2.1 <- sum(scc$dr2.1)
```
**The SCC is $71/ton.**

```{r}
ggplot(data = scc_table, aes(x = dr, y = scc_vector)) +
    geom_line() +
  labs(x = "Discount rate (%)", y = "Social cost of carbon ($/ton)") +
  theme_minimal() +
geom_point(aes(y = scc2.1, x = r), size = 7, shape = 3) +
  geom_text(aes(y = scc2.1, x = 4, label = "SCC = $71/ton at a discount rate of 2.1%."), size = 5, shape = 8)
```


# 5. Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual and Policy B is to take immediate and strong action on climate change. Use these facts

• If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

In principle, we would like the “pulse” to be a single ton of carbon. But this pulse hardly shows up unless you measure temperature out to 10 decimal points. So we introduce a larger one-time pulse and you can infer the “per-ton” effects by dividing the effect by 35 billion.

• Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever.
• Society is risk neutral
• Use a discount rate of 2%

```{r}
policy <- mutated_df %>% 
  mutate(warming_policy_a = 1.5*dam_baseline) %>% # 0.5 chance of this occurring under A
  mutate(warming_policy_b = ifelse(year < 2050, warming_policy_a, 1.29)) %>% # 100% chance of this occurring under B
  mutate(damages_a = dam_function(warming_policy_a)) %>% 
  mutate(damages_b = dam_function(warming_policy_b)) %>%
  mutate(years_into_future = year - 2022) %>% 
  mutate(damages_a_discounted = damages_a/(1.02^years_into_future)) %>%
  mutate(damages_b_discounted = damages_b/(1.02^years_into_future)) %>% 
  mutate(damages_baseline_discounted = dam_baseline/(1.02^years_into_future))
```

# calculate the probability
```{r}
# expected values
a_present_value <- 0.5*sum(policy$damages_baseline_discounted) + 0.5*sum(policy$damages_a_discounted)
b_present_value <- sum(policy$damages_b_discounted)
```

## What is the expected present value of damages up to 2100 under Policy A? 
Expected present value of damages up to 2100 under **Policy A**: $`r round(a_present_value/1e+12, 0)` trillion

## What is the expected present value of damages up to 2100 under Policy B? 
Expected present value of damages up to 2100 under **Policy B**: $`r round(b_present_value/1e+12, 0)` trillion

## Suppose undertaking Policy A costs zero and undertaking Policy B costs X. The total cost of a policy is the implementation cost plus expected damages. Make a plot of total cost of Policy A and total cost of Policy B against X. 

```{r}
a_cost <- function(x){a_present_value}
b_cost <- function(x){b_present_value + x}

potential_cost <- c(0, 1e15, 2e15, 3e15, 4e15, 5e15, 6e15, 7e15, 8e15, 9e15)
cost_df <- data.frame(potential_cost)

total_cost <- cost_df %>%
  mutate(policyA_cost = a_cost(cost_df)) %>% 
  mutate(policyB_cost = b_cost(cost_df))
 
ggplot() +
  geom_hline(yintercept = b_present_value, color = "cornflowerblue") +
  stat_function(fun = b_cost, color = "orange") +
  scale_color_manual(values = c("cornflowerblue", "orange")) +
  ylim(0, 6e+15) +
  xlim(0, 5e+15)
```

## Suppose Congress wants to undertake the policy with the lower total cost. Use your graph to advise Congress.

# 6. Now suppose that society is risk averse. How do you think this would change your advice to Congress?



